---
title: "Quarto_Homework_Task"
author: "Rahul Babu"
format:
  pdf:
    code-block: false
execute:
  echo: false
editor: visual
---

# The MTCars Dataset

## Description:

The mtcars data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).

### Step 1: Load the Mass package and view first few entries

```{r}
#| label: Loader
#| warning: False
#| message: False
library(MASS)
library(ggplot2)
library(knitr)
library(dplyr)

head(mtcars) |>
  kable(caption = "Table 1: First 6 Rows of the mtcars Dataset")
```

From the tabulation of the first few entries of the `mtcars` dataset, we can make the following observations:

-   Each row corresponds to a different car model and includes specifications such as miles per gallon, number of cylinders, horsepower, weight, and transmission type.

-   These variables provide a basis for analyzing automobile performance characteristics of cars from the 1970s.

### Step 2: Display a scatterplot of horsepower vs mpg with respect to cylinders

```{r}
#| label: Plot_Block
ggplot(mtcars, aes(x = hp, y = mpg, color = factor(cyl))) +
  geom_point(size = 3) +
  labs(
    title = "Horsepower vs. MPG in mtcars Dataset",
    x = "Horsepower (hp)",
    y = "Miles Per Gallon (mpg)",
    color = "Cylinders"
  ) +
  theme_minimal()
```

From the scatter-plot, we can make the following observations:

-   A clear negative trend is visible. Cars with higher horsepower tend to have lower miles per gallon.

-   Vehicles with 8 cylinders cluster in the high horsepower, low mpg range, while 4-cylinder cars dominate the high-efficiency, lower horsepower zone.

### Example from MTCars documentation:

```{r}
#| label: Example_From_mtcars

require(graphics)
pairs(mtcars, main = "mtcars data", gap = 1/4)
coplot(mpg ~ disp | as.factor(cyl), data = mtcars,
       panel = panel.smooth, rows = 1)
## possibly more meaningful, e.g., for summary() or bivariate plots:
mtcars2 <- within(mtcars, {
   vs <- factor(vs, labels = c("V", "S"))
   am <- factor(am, labels = c("automatic", "manual"))
   cyl  <- ordered(cyl)
   gear <- ordered(gear)
   carb <- ordered(carb)
})
summary(mtcars2)
```
